---
title: "Solutions for Lab 4 exercises - Distributions II"
author: "Marwin do Carmo"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE)
```

```{r}
library(knitr)
```

1. Estimate the letter occurrence probabilities of all 26 letters in English by measuring a paragraph of English text from wikipedia. (hint use strsplit() to split a paragraph into individual letters) (1 point).

```{r}
my_paragraph <- "In probability theory and statistics, covariance is a measure of the joint variability of two random variables.If the greater values of one variable mainly correspond with the greater values of the other variable, and the same holds for the lesser values (that is, the variables tend to show similar behavior), the covariance is positive. In the opposite case, when the greater values of one variable mainly correspond to the lesser values of the other, (that is, the variables tend to show opposite behavior), the covariance is negative. The sign of the covariance therefore shows the tendency in the linear relationship between the variables. The magnitude of the covariance is not easy to interpret because it is not normalized and hence depends on the magnitudes of the variables. The normalized version of the covariance, the correlation coefficient, however, shows by its magnitude the strength of the linear relation."
the_letters <- tolower(unlist(strsplit(my_paragraph, split="")))

```

```{r}
only_letters <- the_letters[the_letters %in% letters]
letter_count <- as.data.frame(table(only_letters))
```

```{r}

letters_df <- data.frame(
  letras = letters
)

letter_freq <- merge(letters_df, letter_count, by.x = "letras", by.y = "only_letters", all.x = TRUE)

letter_freq[is.na(letter_freq)] <- 0
letter_freq["probabilites"] <- letter_freq["Freq"]/sum(letter_freq["Freq"])

knitr::kable(letter_freq)

```

2. Generate “random” strings of letters that are sampled from a distribution where letter occurrence probability is the same as natural English. Use the probabilities for each letter from this wikipedia article, or use your own estimates from the previous question (2 points).

```{r}
random_words <- c()
for (i in 1:100) {
  random_words[i] = paste0(
    sample(letter_freq$letras, 5, replace = TRUE, 
                      prob = letter_freq$probabilites), 
    collapse = "")
}
```

3. Generate a random walk of 10,000 steps. In a random walk, you are simulating the process of randomly taking a step up or down, as if you are on an infinite staircase. At each step you flip a coin. If you get heads you go up one step, if you get tails you go down one step. Start on step 0, then simulate a random walk for 10,000 steps. Your vector should preserve the step number for each step. For example, if the the first three steps were all heads, then the vector would begin with 0,1,2,3, which indicates a single step up each time. Plot the first 1,000 steps. (1 point)


